Antibiotic seasonalities

### 2017-05-10 temporal/seasonal-traces/drug_seasonality.Rmd
improved seasonality

@@

@@

@@

```{r spectrogram functions}
strongest_period = function(pg) 1.0 / (pg$freq[which.max(pg$spec)])
drug_data = function(drugs) {
  filter(dat, antibiotic %in% drugs) %>%
    group_by(week) %>%
    summarize(n_claims=sum(n_claims))
}

drug_plot = function(drugs) {
  drug_data(drugs) %>%
    ggplot(aes(x=week, y=n_claims)) + geom_point()
}

drug_spectrogram = function(drugs) drug_data(drugs) %$% spec.pgram(n_claims)
drug_period = function(drugs) strongest_period(drug_spectrogram(drugs))
```

@@

Azithromycin and quinolones are strongly seasonal. Tetracyclines are moderately seasonal. Penicillins and ceftriaxone are not seasonal.

@@

```{r azithromycin}
drugs = c('azithromycin')
drug_plot(drugs)
plot(drug_spectrogram(drugs))
```

The strongest period is `r drug_period(drugs)` weeks.

@@

```{r quinolone}
drugs = c('levofloxacin', 'ciprofloxacin')
drug_plot(drugs)
plot(drug_spectrogram(drugs))
```

@@

The strongest period is `r drug_period(drugs)` weeks.

```{r tetra}
drugs = c('tetracycline', 'doxycycline')
drug_plot(drugs)
plot(drug_spectrogram(drugs))
```

The strongest period is `r drug_period(drugs)` weeks.

@@

```{r ceftriaxone}
drugs = c('ceftriaxone')
drug_plot(drugs)
plot(drug_spectrogram(drugs))
```

The strongest period is `r drug_period(drugs)` weeks.

@@

```{r penicillin}
drugs = c('penicillin', 'ampicillin')
drug_plot(drugs)
plot(drug_spectrogram(drugs))
```

The strongest period is `r drug_period(drugs)` weeks.

>>>>

## Data

@@

## Analysis

@@

I also use `nls` to fit the curves. The rough equation is $y = A + Bt + C\sin(2\pi t/D + E\pi)$. The time $t$ is just weeks, and $y$ is the number of claims in that week. This function works only if you give good starting values:

- vertical offset $A$ as the mean number of claims
- secular trend slope $B \approx 0$
- amplitude $C$ as the standard deviation of the number of claims
- period $D \approx 52$ (because we expect a yearly period)
- phase offset $E \approx \tfrac{1}{2}$ (because we expect consumption to be high in the winter when $t \approx 0$)

For the plots, I show a fit that includes a linear

@@

```{r spectrogram_functions, echo=TRUE}
strongest_period = function(pg) 1.0 / (pg$freq[which.max(pg$spec)])
drug_data = function(drugs) {
  filter(dat, antibiotic %in% drugs) %>%
    group_by(week) %>%
    summarize(n_claims=sum(n_claims))
}

drug_plot = function(drugs) {
  pf = periodic_fit(drugs)
  p = pf$m$getPars()
  f = function(t) p['A'] + p['B'] * t + p['C'] * sin(2 * pi * t / p['D'] + pi * p['E'])
  drug_data(drugs) %>%
    ggplot(aes(x=week, y=n_claims)) + geom_point() + stat_function(fun=f)
}

drug_spectrogram = function(drugs) drug_data(drugs) %$% spec.pgram(n_claims)
drug_period = function(drugs) strongest_period(drug_spectrogram(drugs))

periodic_fit = function(drugs) {
  d = drug_data(drugs) %>% mutate(x=week, y=n_claims)
  nls(y ~ A + B*x + C*sin(2*pi*x/D + pi*E), start=list(A=mean(d$y), B=0, C=sd(d$y), D=52, E=0.5), data=d)
}

periodic_fit_table = function(drugs) periodic_fit(drugs) %>% tidy %>% select(term, estimate) %>% kable
```

@@

Azithromycin and quinolones are strongly seasonal: the spectrogram has a clear peak at approximately 52 week period, and the curve fit gives a strong amplitude and a period of almost exaclty 52. The quinolones are also seasonal---there is a peak in the spectrogram and the curve fit works nicely---but the amplitude is not as strong.

The tetracyclines are moderately seasonal: the spectrogram and fit both work, but the amplitudes are again smaller. Ceftriaxone is weakly seasonal: the spectrogram does find that approximately 52 weeks is the strongest period, but the curve fit surprises me. The penicillins are not seasonal at all.

@@

```{r azithromycin}
drugs = c('azithromycin')
drug_plot(drugs)
plot(drug_spectrogram(drugs))
periodic_fit_table(drugs)
```

The strongest period in the spectrogram is `r drug_period(drugs)` weeks.

@@

```{r quinolone}
drugs = c('levofloxacin', 'ciprofloxacin')
drug_plot(drugs)
plot(drug_spectrogram(drugs))
periodic_fit_table(drugs)
```

The strongest period in the spectrogram is `r drug_period(drugs)` weeks.

@@

```{r tetra}
drugs = c('tetracycline', 'doxycycline')
drug_plot(drugs)
plot(drug_spectrogram(drugs))
periodic_fit_table(drugs)
```

The strongest period in the spectrogram is `r drug_period(drugs)` weeks.

@@

```{r ceftriaxone}
drugs = c('ceftriaxone')
drug_plot(drugs)
plot(drug_spectrogram(drugs))
periodic_fit_table(drugs)
```

The strongest period in the spectrogram is `r drug_period(drugs)` weeks.

@@

```{r penicillin}
drugs = c('penicillin', 'ampicillin')
drug_plot(drugs)
plot(drug_spectrogram(drugs))
periodic_fit_table(drugs)
```

The strongest period in the spectrogram is `r drug_period(drugs)` weeks.
